---
title: "Chapter2 CRD for one factor"
output: html_notebook
---

## Example 1.   
```{r}
library(daewr)
# example data set 
bread
```
To generate DoE with randomization, use 'sample'

factor : time 
level : descrete, ordered
```{r}
set.seed(2020)
f <- factor(rep(c(35,40,45),4))
fac <- sample(f,12)
plan <- data.frame(loaf = 1:12, time = fac)

```
 
fit linear model with lm function.

```{r}
mod0 <- lm(height ~ time, data = bread)
summary(mod0)
plot(bread$time,bread$height, type = "p")
```

contrast : '対比'
Estimable contrasts  --> for what ??
```{r}
library(gmodels)
fit.contrast(mod0, 'time', c(1,-1,0))
```

ANOVA 
```{r}
mod1 <- aov(height ~ time, data = bread)
summary(mod1)
```

F value = mean squares of Treatment / mean squares of Error
P value is the probability of exceeding the calcuated F value if the null hypothesis is true.

### Verufying Assumptions of LM

  1. constancy of variance of errors
  2. normality of the errors 
  
```{r}
par(mfrow = c(2,2))
plot(mod1)

plot(residuals(mod1) ~ loaf,
     main = "Residuals vs Exp.Unit",
     font.main = 1,
     data = bread)
abline( h = 0, lty = 2)
```

Residuals vs Factor levels :
  all factor have similar variability of the residuals ? -> if not, variance of the error is not constant. with only 4 replicats in each cell this is difficult to say.
in this case, transformation would be effective.

If the spread in the residuals tends to increase proportionally as a function of the cell means,transformation  Y = f(y) can uually become more sensitive analysis.
Boxcox transformation Y = y^-lambda usually work well.
appropriate lambda can be calculated with MASS pacakge.
```{r}
library(MASS)
bc <- boxcox(mod0)
lambda <- bc$x[which.max(bc$y)]
```
using lambda, which minimize ssE, transform bread data.

```{r}
tbread <- transform(bread, theight = height ^ lambda)
tbread
mod2 <- aov(theight ~ time, data = tbread)
summary(mod2)
par(mfrow = c(2,2))
plot(mod2)
```

