---
title: "Chapter2 CRD for one factor"
output: html_notebook
---

## Example 1.   
```{r}
library(daewr)
# example data set 
bread
```
To generate DoE with randomization, use 'sample'

factor : time 
level : descrete, ordered
```{r}
set.seed(2020)
f <- factor(rep(c(35,40,45),4))
fac <- sample(f,12)
plan <- data.frame(loaf = 1:12, time = fac)

```
 
fit linear model with lm function.

```{r}
mod0 <- lm(height ~ time, data = bread)
summary(mod0)
plot(bread$time,bread$height, type = "p")
```

contrast : '対比'
Estimable contrasts  --> for what ??
```{r}
library(gmodels)
fit.contrast(mod0, 'time', c(1,-1,0))
```

ANOVA 
```{r}
mod1 <- aov(height ~ time, data = bread)
summary(mod1)
```

F value = mean squares of Treatment / mean squares of Error
P value is the probability of exceeding the calcuated F value if the null hypothesis is true.

### Verufying Assumptions of LM

  1. constancy of variance of errors
  2. normality of the errors 
  
```{r}
par(mfrow = c(2,2))
plot(mod1)

plot(residuals(mod1) ~ loaf,
     main = "Residuals vs Exp.Unit",
     font.main = 1,
     data = bread)
abline( h = 0, lty = 2)
```

Residuals vs Factor levels :
  all factor have similar variability of the residuals ? -> if not, variance of the error is not constant. with only 4 replicats in each cell this is difficult to say.
in this case, transformation would be effective.

If the spread in the residuals tends to increase proportionally as a function of the cell means,transformation  Y = f(y) can uually become more sensitive analysis.
Boxcox transformation Y = y^-lambda usually work well.
appropriate lambda can be calculated with MASS pacakge.
```{r}
library(MASS)
bc <- boxcox(mod0)
lambda <- bc$x[which.max(bc$y)]
```
using lambda, which minimize ssE, transform bread data.

```{r}
tbread <- transform(bread, theight = height ^ lambda)
tbread
mod2 <- aov(theight ~ time, data = tbread)
summary(mod2)
par(mfrow = c(2,2))
plot(mod2)
```

When the variance of the experimental error is not constant for all levels of the treatment factor, but it is not related to the cell means, a transformation will not be an appropriate way of equalizing or stabilizing the variances.
分散がFactor level と関係なしにばらつくとき、transformは効果がない。
--> Use weighted least squares.

```{r}
bread
with(bread,  {
  std <- tapply(height, time, sd)
  weights <- rep(1/std, each = 4)
  mod3 <- lm(height ~ time, weights = weights, data = bread)
  anova(mod3)
  }
)
```

When error distribution in not notmal, an alternative is to use a generalized linear model (GLM)
エラーがnormalでないとき、一般化線形モデルを使うのもよい。


## Example 2 :  LM vs GLM
GLMはLMに加えてerror distribution とlink functionを使う。
error distribution (normal in lm) can be poasson, gannma, multinominal distribution
link function (identity in lm) can be Y = log(y), Y = 1/y , etc.

```{r}
library(daewr)
#data set
head(teach)
nrow(teach)

library(MASS)
# orderd logistic or probit regression fucntion : polr

# full model (includes the treatment factor, method)
modf <- polr(score ~ method, weight = count, data = teach)
# reduced model (includes only intercept)
modr <- polr(score ~ 1, weight = count , data = teach)

anova(modf, modr)
```

polr fuction uses logistic link function and multinominal distribution by default.
P = 0.008 for the likelihood ratio chi-square statistic means there is a significant difference of 'method'


### Determining reprecation number 
Define significance level alpha (0.05 would be suitable), and 'practical significance = Delta 効果量'.
'Power' = 1 -beta (probability of type 2 error = False Negative!  ) between 80% and 90% is usually sufficient for most experimental study

Example:
realistic number of replicates = 2~6 
Delta = 3
sigma (experimental error) = sqrt = 2.1

```{r}
bread
alpha <-rep(0.05, 6)
Fpower1(alpha, nlev = 3, nreps = 2:7,  Delta = 3, sigma = sqrt(2.1))
```

With rep = 5 there would be a 73% chance of detecting a difference in cell means as large as 3 (Delta).


## Example 3 : after F-test

```{r}
daewr::sugarbeet

mod4_lm <- lm(yield ~ treat, data = sugarbeet)
summary(mod4_lm)

mod4 <- aov(yield ~ treat, data = sugarbeet)
summary(mod4)
```




